Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-05-07 code/rmarkdown/article.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

## Results
<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species. The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species. Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k,ma.rm=TRUE)`; mean $K$ `r mean(adaptive.k,na.rm=TRUE)`; max $K$ `r max(adaptive.k,na.rm=TRUE)`; Table S2). On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg. _Phyteuma hemisphaericum_; S49). Additionally, most species show east-west patterns in their intra-specific genetic variation (eg. _Dryas octopetala_; S36).

<!-- correlation analysis -->
The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics. After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.
The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated;_P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics. After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{`r env.surrogacy.AOV$Resid..Dev[[4]]`} P = `r env.surrogacy.AOV$Pr..Chi.[[4]]`$). The slope for each species were all significantly different to zero (minimum $z value = `r min(env.surrogate.spp.DF$statistic)`$; maximum $P = r `max(env.surrogate.spp.DF$p.value)`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r mean(env.surrogate.spp.DF$r2CU)` $R^2$ `r sd(env.surrogate.spp.DF$r2CU)` SD; Figure S56), with exception to _Carex sempervirens_. The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets. Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{`r geo.surrogacy.AOV$Resid..Dev[[4]]`} P = `r geo.surrogacy.AOV$Pr..Chi.[[4]]`$). The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r mean(geo.surrogate.spp.DF$r2CU)` $R^2$ `r sd(geo.surrogate.spp.DF$r2CU)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species. Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 3). The prioritizations generated under the single-species scenario showed relatively homogeneous selection frequencies across the study area--regardless of the combination of targets used to generate them (Figures 3a, 3b, 3c). On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them. The prioritization generated using only amount-based targets contained fewer planning units (Figure 3d; $n = `r sum(selections(multi.spp.prioritisations[[1]]))`$) than the prioritizations generated using additional of surrogate-based targets (Figure 3e; $n = `r sum(selections(multi.spp.prioritisations[[2]]))`$) or genetic-based targets (Figure 3f; $n = `r sum(selections(multi.spp.prioritisations[[3]]))`$). The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios. These prioritizations tended to concentrate conservation effort towards areas with higher elevation. Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations ((Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$)).

@@

The prioritizations generated using additional surrogate-based targets met the neutral genetic variation targets for more species than the prioritizations generated using just amount-based targets (Figure 3). The biggest difference between the prioritizations generated using additional surrogate-based targets and just the amount-based targets was in the single-species scenarios (`r ` % and `r ` % of species meeting genetic variation targets respectively). In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets. Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila_repens_ and _Luzula_alpinopilosa_). These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.

>>>>

## Results
<!-- summary of methods -->
We successfully mapped the main spatial patterns of adaptive (if present) and neutral variation for each of the 26 species. The outlier detection methods detected adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` species. Generally, two outlier detection methods gave fairly consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (all species $K = 2$; Table S2). On the other hand, a few more continuous dimensions were often needed for neutral genetic variation for each species (minimum $K = `r min(neutral.k)`$; mean $K = `r mean(neutral.k)`$; max $K = `r max(neutral.k)`$; Table S2). The ordinations resulting from the NMDS analyses were used to construct an adaptive (if present) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally show patterns of spatial autocorrelation--especially the adaptive genetic spaces (eg. _Phyteuma hemisphaericum_; S49). Additionally, most species show east-west patterns in their intra-specific genetic variation (eg. _Dryas octopetala_; S36).

<!-- correlation analysis -->
The dissimilarity matrices generated using the surrogate variables correlated with the dissimilarity matrices based on the genetic characteristics of individuals for most species (Table 1). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species (`r round((sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)/length(na.omit(correlation.model.summary.DF$Environmental.corr.P)))*100,2)` % associated with loci under selection; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species (`r round((sum(correlation.model.summary.DF$Geographic.corr.P<0.05)/length(na.omit(correlation.model.summary.DF$Geographic.corr.P)))*100,2)` % of all the species investigated; _P_ < 0.05; mean marginal `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics. After verifying that the geographic and environmental variables correlate with genetic variation, the next step was to determine how effective these variables were as surrogates in the context of conservation planning.

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion environment variation it also secured differed depending on the species for which the prioritization was generated ($\chi^2_{r env.surrogacy.AOV$df[[4]]} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`} P = pvalString(`r env.surrogacy.AOV$Pr..Chi.[[4]]`)$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(min(env.surrogate.spp.DF$statistic),2)`; maximum $P = `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56), with exception to _Carex sempervirens_. The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets. Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization ($\chi^2_{r geo.surrogacy.AOV$df[[4]]} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum $z value = `r min(geo.surrogate.spp.DF$statistic)`$; maximum $P = `r max(env.surrogate.spp.DF$p.value)`; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species. Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 3). The prioritizations generated under the single-species scenario showed relatively homogeneous selection frequencies across the study area--regardless of the combination of targets used to generate them (Figures 3a, 3b, 3c). On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them. The prioritization generated using only amount-based targets contained fewer planning units (Figure 3d; $n = `r sum(selections(multi.spp.prioritisations[[1]]))`$) than the prioritizations generated using additional of surrogate-based targets (Figure 3e; $n = `r sum(selections(multi.spp.prioritisations[[2]]))`$) or genetic-based targets (Figure 3f; $n = `r sum(selections(multi.spp.prioritisations[[3]]))`$). The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios. These prioritizations tended to concentrate conservation effort towards areas with higher elevation. Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations (Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$).

@@

The prioritizations generated using additional surrogate-based targets met the neutral genetic variation targets for more species than the prioritizations generated using just amount-based targets (Figure 3). The biggest difference between the prioritizations generated using additional surrogate-based targets and just the amount-based targets was in the single-species scenarios (`r ` % and `r ` % of species meeting genetic variation targets respectively). In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets. Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.

### 2016-05-07 code/rmarkdown/supporting_information.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

calc.fig.height <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	return((((n.plots-1) %/% 4)+1) * 1)
}

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures. Points represent values associated with randomly generated prioritizations. Blue lines indicate average trends computed using a generalised linear model with a logit link. Each panel shows data for a different species.'}
# format data
curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

@@

```{r, echo=FALSE, results='asis'}
adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=pvalString(Stress))
neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=pvalString(Stress))
names(adaptive.spp.nmds.summary.DF)[2:5] <- paste0('adaptive.', names(adaptive.spp.nmds.summary.DF)[2:5])
names(neutral.spp.nmds.summary.DF)[2:5] <- paste0('neutral.', names(neutral.spp.nmds.summary.DF)[2:5])
spp.nmds.summary.DF2 <- full_join(adaptive.spp.nmds.summary.DF, neutral.spp.nmds.summary.DF, by='Species')
spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)],'}')
spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)],'}')
spp.nmds.summary.DF2 <- spp.nmds.summary.DF2 %>%
	mutate(blank='') %>%
	select(Species, adaptive.k:adaptive.Stress, neutral.k:neutral.Stress)
curr.file <- tempfile(fileext='.tex')
x <- spp.nmds.summary.DF2 %>%
	latex(
		file=curr.file,
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 5),
		cgroup=c('', 'Adaptive', 'Neutral'), n.cgroup=c(1, 2, 2),
		colheads=c('\\bfseries{Species}', 'K', 'Stress', 'K', 'Stress'), here=TRUE,
		caption='Summary of non-metric multi-dimensional scaling (NMDS) analyses on genetic variation for each species. The analyses were used to identify the main graondients of variation in binary adaptive and neutral loci data. Bold stress values indicate NMDS analyses that have converged.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

>>>>

calc.fig.height <- function(i) {
	# init
	adapt.col <- paste0(unique(spp.samples.DF$species)[i], '_adaptive_d1')
	neutral.col <- paste0(unique(spp.samples.DF$species)[i], '_neutral_d1')
	n.plots <- 0 
	# count adaptive plot
	if (adapt.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['adaptive']]$ndim
	# count neutral plot
	if (neutral.col %in% names(grid.FPLY))
		n.plots <- n.plots + spp.nmds.LST[[i]][['neutral']]$ndim
	# exports
	return((((n.plots-1) %/% 4)+1) * 2)
}

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=2.5, fig.cap='The relationship between the proportion of adaptive genetic variation secured in a prioritizaiton and the proportion of environmental variation it also secures. Points represent values associated with randomly generated prioritizations. Blue lines indicate average trends computed using a generalised linear model with a logit link. Each panel shows data for a different species.'}
# format data
curr.DF <- env.surrogacy.DF %>% filter(!is.na(genetic.held)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2 <- env.surrogate.spp.DF %>% filter(!is.na(r2CU)) %>% mutate(Species=gsub('_', ' ', Species))
curr.DF2$R2_repr <- round(curr.DF2$r2CU,2)
curr.DF2$R2_repr <- sapply(curr.DF2$R2_repr, function(x) {as.character(as.expression(substitute(~~italic(R)^2~"="~r2, list(r2=x))))})
curr.DF2$surrogate.held <- 0
curr.DF2$genetic.held <- 0.9

@@

```{r, echo=FALSE, results='asis'}
adaptive.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='adaptive') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
neutral.spp.nmds.summary.DF <- spp.nmds.summary.DF %>% filter(Loci=='neutral') %>% mutate(Stress=as.character(round(Stress,3))) %>% mutate(Stress=replace(Stress, Stress=='0', '< 0.001'))
names(adaptive.spp.nmds.summary.DF)[2:5] <- paste0('adaptive.', names(adaptive.spp.nmds.summary.DF)[2:5])
names(neutral.spp.nmds.summary.DF)[2:5] <- paste0('neutral.', names(neutral.spp.nmds.summary.DF)[2:5])
spp.nmds.summary.DF2 <- full_join(adaptive.spp.nmds.summary.DF, neutral.spp.nmds.summary.DF, by='Species')
spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$adaptive.Stress[which(spp.nmds.summary.DF2$adaptive.Converged)],'}')
spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)] <- paste0('\\bfseries{', spp.nmds.summary.DF2$neutral.Stress[which(spp.nmds.summary.DF2$neutral.Converged)],'}')
spp.nmds.summary.DF2 <- spp.nmds.summary.DF2 %>%
	mutate(blank='') %>%
	select(Species, adaptive.k:adaptive.Stress, neutral.k:neutral.Stress)
curr.file <- tempfile(fileext='.tex')
x <- spp.nmds.summary.DF2 %>%
	latex(
		file=curr.file,
		rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, col.just=rep('c', 5),
		cgroup=c('', 'Adaptive', 'Neutral'), n.cgroup=c(1, 2, 2),
		colheads=c('\\bfseries{Species}', 'K', 'Stress', 'K', 'Stress'), here=TRUE,
		caption='Summary of non-metric multi-dimensional scaling (NMDS) analyses on genetic variation for each species. The analyses were used to identify the main graondients of variation in binary adaptive and neutral loci data. Bold stress values indicate NMDS analyses that have converged.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{3-4} \\cline{6-', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
```

### 2016-05-07 code/rmarkdown/tables.Rmd
+ fixes to article
Former-commit-id: b60c0fbeff1f1c52aff2c2fee15295295b8f438c

```{r, echo=FALSE, results='asis'}
curr.file <- tempfile(fileext='.tex')
x <- correlation.model.summary.DF %>%
	select(-contains('raw')) %>%
	mutate(
		Environmental.corr.P=pvalString(Environmental.corr.P),
		Geographic.corr.P=pvalString(Geographic.corr.P), 
		blank=rep('', length(Species)),
		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
	latex(file=curr.file, 
		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
		caption='Summary of maximum likelihood population-effects (MLPE) models. Briefly, for each species associated with adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group). Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group). Data shows the marginal $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model, and also the results from comparing the models to their corresponding null models. All $\\chi^2$ tests had 1 degree of freedom.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
 ```

>>>>

```{r, echo=FALSE, results='asis'}
curr.file <- tempfile(fileext='.tex')
x <- correlation.model.summary.DF %>%
	select(-contains('raw')) %>%
	mutate(
		Environmental.corr.P=pvalString(Environmental.corr.P),
		Geographic.corr.P=pvalString(Geographic.corr.P),
		Geographic.Marginal.R2=pvalString(Geographic.Marginal.R2),
		Geographic.Conditional.R2=pvalString(Geographic.Conditional.R2),
		Environmental.Marginal.R2=pvalString(Environmental.Marginal.R2),
		Environmental.Conditional.R2=pvalString(Environmental.Conditional.R2),
		blank=rep('', length(Species)),
		Species=paste0('\\textit{', gsub('\\ ', ' ', Species), '}')) %>%
	select(Species:Environmental.corr.P, Geographic.Marginal.R2:Geographic.corr.P) %>%
	latex(file=curr.file, 
		cgroup=c('', 'Environmental', 'Geographic'), n.cgroup=c(1, 4, 4), col.just=rep('c', 9),
		digits=2, rowname=NULL, first.hline.double=FALSE, booktabs=TRUE, here=TRUE,
		colheads=c('\\bfseries{Species}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}', '$MR^2$', '$CR^2$', '$\\chi^2$', '\\textit{P}'),
		caption='Summary of maximum likelihood population-effects (MLPE) models. Briefly, for each species associated with adaptive loci, a model was fit to correlate dissimilarities among planning units where the species in terms of the adaptive genetic characteristics of individuals in them with dissimilarities in terms of the environmental conditions inside the planning units (denoted in the environmental column group). Similarly, for each species, a model was fit to correlate dissimilarities among the planning units in terms of the neutral genetic characteristics of indiviudals in them with dissimilarities in terms of the geographic position of the planning units (denoted in the geographic column group). Data shows the marginal $R^2$ ($MR^2$) and conditional $R^2$ ($CR^2$) for each model, and also the results from comparing the models to their corresponding null models. All $\\chi^2$ tests had 1 degree of freedom.'
	)
readLines(curr.file)[-1] %>% paste0(collapse='\n') %>% 
	gsub(pattern='\\cline{1-', replacement='\\cline{2-6} \\cline{8-', fixed=TRUE) %>% 
	gsub(pattern='{c}{\\textit{P}}\\tabularnewline', replacement='{c}{\\textit{P}}\\T\\B\\tabularnewline', fixed=TRUE) %>% 
	gsub(pattern='toprule', replacement='toprule[1pt]', fixed=TRUE) %>%
	gsub(pattern='bottomrule', replacement='bottomrule[1pt]', fixed=TRUE) %>%
	cat()
 ```
